M.Sc. Proposal: \\ \bigskip Computational Electrokinetics

### 2012-12-25 notes/thesis-latex-template/thesis.tex
Thesis

\subsection{Computational Grid}
A regular grid of size $n_r \times n_\theta$ is defined, with:
\begin{align} \label{eq:grid}
\nonumber
(r_i,\theta_j) &\in [1, \infty) \times [0,\pi], \\ 
r_i &= (1+\Delta_r)^i, \\
\nonumber
\theta_j &= \Delta_\theta \cdot j,
\end{align}
where a logarithmic grid spacing has been chosen for $r$ and a uniform grid for $\theta$.
Note that $r_0 = 1$ and $R_{max} = (1+\Delta_r)^{n_r}$, where
$\Delta_r = \pars{R_{max}} ^ {1/n_r} - 1$ and $\Delta_\theta = {\pi/n_\theta}$.
Because the grid is finite, the choice 
$R_{max}$ must be taken large enough so as to have a negligible effect
on the solution.
This grid induces a disjoint subdivision of the domain 
$\Omega = [1, R_{max}] \times [0,\pi] = \bigcup_{ij}\Omega_{ij}$ into cells.
A specific cell $\Omega_{ij}$ and its center $(\bar{r}_i, \bar{\theta}_j)$ are defined by \eqref{eq:grid}:
\begin{align}
\nonumber
\Omega_{ij} &= [r_{i-1}, r_{i}] \times [\theta_{j-1}, \theta_{j}], \\
\bar{r}_i &= (r_{i-1} + r_{i})/2, \\
\nonumber
\bar{\theta}_j &= (\theta_{j-1} + \theta_{j})/2.
\end{align}

Each discrete variable is located with respect of its cell 
as shown at Figure \ref{fig:grids}.
$\varPhi$, $C$ and $P$ are represented by their value at the center of each cell, 
using an all-centered grid, whereas $\bV$ is represented by its values at cell 
boundaries, using a staggered grid:
\begin{align}
\nonumber \varPhi^h_{[i,j]} &= \varPhi(\bar{r}_i, \bar{\theta}_j), \\
\nonumber C^h_{[i,j]} &= C(\bar{r}_i, \bar{\theta}_j), \\
P^h_{[i,j]} &= P(\bar{r}_i, \bar{\theta}_j), \\
\nonumber V_r^h{}_{[i,j]} &= V_r(r_i, \bar{\theta}_j), \\
\nonumber V_\theta^h{}_{[i,j]} &= V_\theta(\bar{r}_i, {\theta}_j).
\end{align}

\begin{figure}[htbp]
    \begin{center}
	\includegraphics[width=0.8\textwidth]{figs/grids.eps}
        \caption[Computational grid]{The computational grid is defined using
        a spherical coordinate system, due to axisymmetry of the physical problem
        and boundary conditions discretization on the particle surface.}
    \label{fig:grids}
    \end{center}
\end{figure}

@@

\section{Source code overview}

\subsection{Variable}
We define the variable vector $\bx$ as \verb|Variable| 
operator using the following MATLAB code:

\begin{lstlisting}
classdef Variable < handle
properties
    value;
end
methods
    function self = Variable(varargin)
        self.value = col(varargin{:});
    end
    function r = res(self)
        r = self.value(:);
    end
    function G = grad(self) 
        G = speye(numel(self.value));
    end
    function update(self, dv)
        self.value = self.value + dv;
    end
end
\end{lstlisting}
Then the initial solution for the solver is set using
$\bx = \verb|Variable|(\varPhi, C, V_r, V_t, P)$
from the initial conditions of the solver.

\subsection{Grid}
Most of the operators are defined a \verb|Grid| object,
enabling validation of operators compatibility and 
convenient interpolation between grids. Note that
each operator must define its \verb|res()| and \verb|grad()|
functions, for residual and gradient computation, respectively.
\begin{lstlisting}
% Regular structured grid class
classdef Grid < handle
properties
    x, y;
    X, Y;
    size;    
    numel;
end
methods
    function self = Grid(x, y)
        self.x = x(:);
        self.y = y(:);
        [self.X, self.Y] = ndgrid(self.x, self.y);
        self.size = [numel(self.x), numel(self.y)];
        self.numel = prod(self.size);
    end
end
\end{lstlisting}
\begin{lstlisting}
classdef Operator < handle
properties
    grid; % a grid object for the output
end
methods
    function op = Operator(grid)
        op.grid = grid;
    end
    function op = plus(op1, op2)
        op = Sum(op1, op2);
    end
    function op = minus(op1, op2)
        op = Minus(op1, op2);
    end
    function op = uminus(op)
        op = Product(-1, op);
    end
    function op = mtimes(op1, op2)
        op = Product(op1, op2);
    end
    function op = log(op)
        op = Func(op, 'log(x)');
    end
    function op = exp(op)
        op = Func(op, 'exp(x)');
    end
    function op = sinh(op)
        op = Func(op, 'sinh(x)');
    end
    function op = cosh(op)
        op = Func(op, 'cosh(x)');
    end
    function op = tanh(op)
        op = Func(op, 'tanh(x)');
    end
end
\end{lstlisting}
Thus, various computations can be applied on simple operators, 
creating more complex operators, such as
\verb|op4 = op1 * exp(op2 + tanh(op3))|. 
Since each operator can compute its residual and gradient, we
can compute automatically the residual and gradient of \verb|o4|
from these of \verb|o1|, \verb|o2|, \verb|o3| using the derivation
rules described at \ref{sec:operator-interface}.

\subsection{Operators}
An operator can be created with $n$ existing operators as input.
\verb|Const| is an operator that corresponds to $n=0$, whose 
residual value is constant and gradient matrix is zero.

\begin{lstlisting}
classdef Const < Operator
properties
    val;
end
methods
    function self = Const(grid, val)
        self = self@Operator(grid);
        if isa(val, 'float')
            if numel(val) == 1
                val = repmat(val, grid.size);
            end
            self.val = val;
        else % function handle evaluated point-wise
            if isa(val, 'char')
                val = str2func(['@(r,t) ' val]);
            end
            self.val = arrayfun(val, grid.R, grid.T);
        end
    end
    function r = res(self)
        r = self.val(:);
    end
    function G = grad(self)
        G = sparse(0); % [0]
    end
end
\end{lstlisting}

\subsubsection{Unary operators}
Unary operators have one existing operator as input. 
\verb|Func| operator implements point-wise analytic function
application (whose derivative is computed using symbolic differentiation toolbox) on the same grid of the input operator. 
\verb|Linear| operator is implemented
using multiplication by a constant sparse matrix 
\verb|L|, which is also the gradient of the operator.
\begin{lstlisting}
classdef Func < Operator
properties
    op; % Input operator
    func; % y = f(x)
    deriv; % dy/dx = f'(x)
end
methods
    function self = Func(op, func)
        grid = op.grid; % Reuse operator's grid
        self = self@Operator(grid);
        self.op = op;
        if isa(func, 'char')
        	func = sym(func); 
        end 
        self.func = matlabFunction(func);
        self.deriv = matlabFunction(diff(func));
    end
    function r = res(self)
        r = self.func( self.op.res() );
    end
    function G = grad(self)
        r = self.op.res();
        G = spdiag( self.deriv(r) ) * self.op.grad();
    end
end
\end{lstlisting}
\begin{lstlisting}
classdef Linear < Operator
properties
    L;
    op;
end
methods
    function self = Linear(grid, op, L)
        self = self@Operator(grid);
        self.op = op;
        self.L = L;
    end
    function r = res(self)
        r = self.L * self.op.res();
    end
    function G = grad(self)
        G = self.L * self.op.grad();
    end
end
\end{lstlisting}
\verb|Deriv| and \verb|Interp| operators (implementing discrete
differentiation and grid interpolation) derive from \verb|Linear|
base class operator and specify \verb|L| sparse matrix accordingly.

@@

\subsubsection{$N$-ary operators}
All equations and boundary condition operators are defined
as operators \verb|{op1, op2, ... opN}|. 
Then, they are combined using the $N$-ary operator 
\verb|op = Join(op1, op2, ... opN)|, creating
the system operator \verb|op|.

\begin{lstlisting}
classdef Join < handle
properties
    ops;
end
methods
    function self = Join(varargin)
        self.ops = varargin;
        for k = 1:numel(self.ops),
            op = self.ops{k};
        end
    end
    function r = res(self)
        r = cell(numel(self.ops), 1);
        for k = 1:numel(self.ops)
            r{k} = self.ops{k}.res();
        end
        r = cat(1, r{:});
    end
    function G = grad(self) 
        G = cell(numel(self.ops), 1);
        for k = 1:numel(self.ops)
            G{k} = self.ops{k}.grad();
        end
        G = cat(1, G{:});
    end
end
\end{lstlisting}

\subsection{Sample source code}
The following example shows how to define the following boundary-value problem:
\begin{align}
\Laplacian F(x,y) &= 1 & (x,y) &\in [0,1]^2 \\
F(0, y) = F(1, y) &= 0 & y &\in [0,1] \\
\deriv{F}{y}(x, 0) = \deriv{F}{y}(x, 1) &= 0 & x &\in [0,1]
\end{align}

\begin{lstlisting}
% Grid definition
x = linspace(0, 1);
y = linspace(0, 1);
g = Grid(x, y);
x_ = (x(1:end-1) + x(2:end)) * 0.5;
y_ = (y(1:end-1) + y(2:end)) * 0.5;
xi = x(2:end-1);
yi = y(2:end-1);
% Variable Definition
v = Variable(zeros(g.size));
F = Linear(g, v, speye(g.numel));
% Equations
Fx = Deriv(Grid(x_, y), F, 1);
Fy = Deriv(Grid(x, y_), F, 2);
Fxx = Deriv(Grid(xi, yi), Fx, 1);
Fyy = Deriv(Grid(xi, yi), Fy, 2);
eq = {(Fxx + Fyy) - Const(1)};             % Laplacian(F) = 1
% Boundary conditions
bnd = {Interp(Grid(x(1)  , y(2:end-1)), F); ... % F(0, y) = 0
	   Interp(Grid(x(end), y(2:end-1)), F); ... % F(1, y) = 0
       Interp(Grid(x, y_(1)  ), Fy); ...    % dF/dy(x, 0) = 0
       Interp(Grid(x, y_(end)), Fy)}		% dF/dy(x, 1) = 0
% System operator
ops = cat(1, eq, bnd)
op = Join(ops{:})
% Solve by Newton iteration
r = op.res()
G = op.grad()
v.update(-(G\r))
\end{lstlisting}

@@

For $\beta = 0$ (no electric field is applied), the steady-state solution is $\cU = 0$:
\begin{equation}\begin{array}{cccc}
\varPhi_0(r,\theta) = 0, &
C_0(r,\theta) = 1, &
\bV_0(r,\theta) = \bzero, &
P_0(r,\theta) = 0.
\end{array}\end{equation}
Assuming $\beta \ll 1$, the linearized equations and boundary conditions are
\begin{equation} \begin{array}{ccc}
\Laplacian \varPhi_1 = 0, &
\Laplacian C_1 = 0, &
\bLaplacian \bV_1 - \bnabla P_1 = \bzero.
\end{array}\end{equation}
The first-order terms $\bx_1$ are given by (see \cite{yariv2010migration} for full derivation)
\begin{equation} \begin{array}{cccc}
\varPhi_1 = \pars{\frac{1}{4r^2} - r}\cos\theta, &
C_1 = \frac{3}{4r^2} \cos\theta, &
\Psi_1 = \cU_1 \pars{\frac{1}{r} - r^2} \frac{\sin^2\theta}{2}, &
P_1 = 0,
\end{array} \end{equation}
and the linear velocity term is:
\begin{equation} \label{eq:linear_velocity}
\cU_1(\gamma) = 2 \log \pars{\frac{1 + \gamma^{-\frac{1}{2}}}{2}}.
\end{equation}
For $\gamma > 1$, the Debye layer has a positive space charge, 
corresponding to a negative particle charge -- 
yielding negative drift velocity, $\cU < 0$. 
For $\gamma < 1$, we have $\cU > 0$.

@@

\subsection{Ion-exchanger results}
For $\beta \ll 1$, the linear regime is dominant \eqref{eq:linear_velocity}.
When $\gamma \approx 1$, the cubic regime, where $\cU = O(\beta^3)$, 
dominates the linear one, even for weak electric fields. 
This phenomenon is confirmed both by the numerical
results and by asymptotic expansion of the nonlinear system (for $\alpha=0$).
The switch from the linear regime to the cubic regime occurs at 
an electric field value of $\beta = O(\sqrt{|\gamma - 1|})$.
The numerical results $\cU^h$, 
as well as the analytical ones \eqref{eq:cubic}, are given in Figure \ref{fig:IonExCubic}.
The numerical solver uses $N_r \times N_\theta = (400 \times 50)$, 
$R_{\max} = 10^7$,
$K = 3$, $M = 5$ and $\beta = 10^{t}$ for $t = -2:0.1:0.5$.

@@

Since the leading order of the solution $\cU(\beta)$ is linear for $\beta \ll 1$, 
the numerical discretization error is $O(\beta)$ as well.
In order to reduce the discretization error of the numerical solution \eqref{eq:disc_zero_force}
and after verifying that the discretization error converges as $O(h^2)$,
we employ Richardson extrapolation to $\cU^h$ and $\cU^{2h}$:
\begin{equation}
\hat\cU = \frac{4}{3} \cU^h - \frac{1}{3} \cU^{2h}.
\end{equation}
The results of the extrapolation and $\hat\cU$ the analytical results 
are given at Figures \ref{fig:Electrophoresis1} and \ref{fig:Electrophoresis2}.
The numerical solver uses $R_{\max} = 10^2$,
$K = 3$, $M = 5$ and $\beta = 0.1 \cdot 2^{t}$ for $t = -2:0.5:5$.
The solver is applied to 2 grids of $N_r \times N_\theta = (128 \times 128)$ as the coarse grid,
and $N_r \times N_\theta = (256 \times 256)$ as the fine grid, in order to use Richardson extrapolation for $\hat\cU$.

@@

>>>>

\subsection{Computational Grid}
A regular grid of size $n_r \times n_\theta$ is defined, with:
\begin{align} \label{eq:grid}
\nonumber
(r_i,\theta_j) &\in [1, \infty) \times [0,\pi], \\ 
r_i &= (1+\Delta_r)^i, \\
\nonumber
\theta_j &= \Delta_\theta \cdot j,
\end{align}
where a logarithmic grid spacing has been chosen for $r$ and a uniform grid for $\theta$.
Note that $r_0 = 1$ and $R_{max} = (1+\Delta_r)^{n_r}$, where
$\Delta_r = \pars{R_{max}} ^ {1/n_r} - 1$ and $\Delta_\theta = {\pi/n_\theta}$.

Because the grid is finite, the choice $R_{max}$ must be taken large enough 
so as to have a negligible effect on the solution. 
On the other hand, in order to have high resolution grid near the particle surface,
$\Delta_r$ should be minimized. It should be noted that for given grid size $n_r$,
these requirements cannot be satisfied simultaneously -- thus, a compromise is required
between large $R_{max}$ and small $\Delta_r$.

This grid induces a disjoint subdivision of the domain 
$\Omega = [1, R_{max}] \times [0,\pi] = \bigcup_{ij}\Omega_{ij}$ into cells.
A specific cell $\Omega_{ij}$ and its center $(\bar{r}_i, \bar{\theta}_j)$ are defined by \eqref{eq:grid}:
\begin{align}
\nonumber
\Omega_{ij} &= [r_{i-1}, r_{i}] \times [\theta_{j-1}, \theta_{j}], \\
\bar{r}_i &= (r_{i-1} + r_{i})/2, \\
\nonumber
\bar{\theta}_j &= (\theta_{j-1} + \theta_{j})/2.
\end{align}

Each discrete variable is located with respect of its cell 
as shown at Figure \ref{fig:grids}.
$\varPhi$, $C$ and $P$ are represented by their value at the center of each cell, 
using an all-centered grid, whereas $\bV$ is represented by its values at cell 
boundaries, using a staggered grid:
\begin{align}
\nonumber \varPhi^h_{[i,j]} &= \varPhi(\bar{r}_i, \bar{\theta}_j), \\
\nonumber C^h_{[i,j]} &= C(\bar{r}_i, \bar{\theta}_j), \\
P^h_{[i,j]} &= P(\bar{r}_i, \bar{\theta}_j), \\
\nonumber V_r^h{}_{[i,j]} &= V_r(r_i, \bar{\theta}_j), \\
\nonumber V_\theta^h{}_{[i,j]} &= V_\theta(\bar{r}_i, {\theta}_j).
\end{align}
\begin{figure}[h]
    \begin{center}
	\includegraphics[width=0.8\textwidth]{figs/grids.eps}
        \caption[Computational grid]{The computational grid is defined using
        a spherical coordinate system, due to axisymmetry of the physical problem
        and boundary conditions discretization on the particle surface.}
    \label{fig:grids}
    \end{center}
\end{figure}

@@

\section{Source code overview}
In the following section we present the source code of the numerical framework and demonstrate its
usage for solving a boundary-value problem.
\subsection{Variable}
We define the variable vector $\bx$ as \verb|Variable| 
operator (Figure \ref{fig:variable.m}).

\begin{figure}[htpb]
\begin{lstlisting}
classdef Variable < handle
properties
    value;
end
methods
    function self = Variable(varargin)
        self.value = col(varargin{:});
    end
    function r = res(self)
        r = self.value(:);
    end
    function G = grad(self) 
        G = speye(numel(self.value));
    end
    function update(self, dv)
        self.value = self.value + dv;
    end
end
\end{lstlisting}
\caption{Variable class source code.} \label{fig:variable.m}
\end{figure}
Then the initial solution for the solver is set using
$\bx = \verb|Variable|(\varPhi, C, V_r, V_t, P)$
from the initial conditions of the solver.

\subsection{Grid}
Most of the operators are using a \verb|Grid| object (Figure \ref{fig:grid.m}),
enabling validation of operators compatibility and 
convenient interpolation between grids. Note that
each operator must define its \verb|res()| and \verb|grad()|
functions, for residual and gradient computation, respectively.
An \verb|Operator| interface is used (Figure \ref{fig:operator.m}) 
to support operator overloading MATLAB syntax.
\begin{figure}[h]
\begin{lstlisting}
% Regular structured grid class
classdef Grid < handle
properties
    x, y;
    X, Y;
    size;    
    numel;
end
methods
    function self = Grid(x, y)
        self.x = x(:);
        self.y = y(:);
        [self.X, self.Y] = ndgrid(self.x, self.y);
        self.size = [numel(self.x), numel(self.y)];
        self.numel = prod(self.size);
    end
end
\end{lstlisting}
\caption{Grid class source code.} \label{fig:grid.m}
\end{figure}
\begin{figure}[h]
\begin{lstlisting}
classdef Operator < handle
properties
    grid; % a grid object for the output
end
methods
    function op = Operator(grid)
        op.grid = grid;
    end
    function op = plus(op1, op2)
        op = Sum(op1, op2);
    end
    function op = minus(op1, op2)
        op = Minus(op1, op2);
    end
    function op = uminus(op)
        op = Product(-1, op);
    end
    function op = mtimes(op1, op2)
        op = Product(op1, op2);
    end
    function op = log(op)
        op = Func(op, 'log(x)');
    end
    function op = exp(op)
        op = Func(op, 'exp(x)');
    end
    function op = sinh(op)
        op = Func(op, 'sinh(x)');
    end
    function op = cosh(op)
        op = Func(op, 'cosh(x)');
    end
    function op = tanh(op)
        op = Func(op, 'tanh(x)');
    end
end
\end{lstlisting}
\caption{Operator interface source code.} \label{fig:operator.m}
\end{figure}

Thus, various computations can be applied on simple operators, 
creating more complex operators, such as
\verb|op4 = op1 * exp(op2 + tanh(op3))|. 
Since each operator can compute its residual and gradient, we
can compute automatically the residual and gradient of \verb|o4|
from these of \verb|o1|, \verb|o2|, \verb|o3| using the derivation
rules described at \ref{sec:operator-interface}.

\subsection{Operators}
An operator can be created with $n$ existing operators as input.
\verb|Const| is an operator that corresponds to $n=0$, whose 
residual value is constant and gradient matrix is zero (see Figure \ref{fig:const.m}).

\begin{figure}[h]
\begin{lstlisting}
classdef Const < Operator
properties
    val;
end
methods
    function self = Const(grid, val)
        self = self@Operator(grid);
        if isa(val, 'float')
            if numel(val) == 1
                val = repmat(val, grid.size);
            end
            self.val = val;
        else % function handle evaluated point-wise
            if isa(val, 'char')
                val = str2func(['@(r,t) ' val]);
            end
            self.val = arrayfun(val, grid.R, grid.T);
        end
    end
    function r = res(self)
        r = self.val(:);
    end
    function G = grad(self)
        G = sparse(0); % [0]
    end
end
\end{lstlisting}
\caption{Constant operator source code.} \label{fig:const.m}
\end{figure}

\subsubsection{Unary operators}
Unary operators have one existing operator as input. 
\verb|Func| operator implements point-wise analytic function
application (whose derivative is computed using symbolic differentiation toolbox) on the same grid of the input operator (Figure \ref{fig:func.m}). 
\verb|Linear| operator is implemented
using multiplication by a constant sparse matrix 
\verb|L|, which is also the gradient of the operator
(Figure \ref{fig:linear.m}).
\verb|Deriv| and \verb|Interp| operators (implementing discrete
differentiation and grid interpolation) derive from \verb|Linear|
base class operator and specify \verb|L| sparse matrix accordingly.

\begin{figure}[h]
\begin{lstlisting}
classdef Func < Operator
properties
    op; % Input operator
    func; % y = f(x)
    deriv; % dy/dx = f'(x)
end
methods
    function self = Func(op, func)
        grid = op.grid; % Reuse operator's grid
        self = self@Operator(grid);
        self.op = op;
        if isa(func, 'char')
        	func = sym(func); 
        end 
        self.func = matlabFunction(func);
        self.deriv = matlabFunction(diff(func));
    end
    function r = res(self)
        r = self.func( self.op.res() );
    end
    function G = grad(self)
        r = self.op.res();
        G = spdiag( self.deriv(r) ) * self.op.grad();
    end
end
\end{lstlisting}
\caption{Pointwise function source code.} \label{fig:func.m}
\end{figure}
\begin{figure}[h]
\begin{lstlisting}
classdef Linear < Operator
properties
    L;
    op;
end
methods
    function self = Linear(grid, op, L)
        self = self@Operator(grid);
        self.op = op;
        self.L = L;
    end
    function r = res(self)
        r = self.L * self.op.res();
    end
    function G = grad(self)
        G = self.L * self.op.grad();
    end
end
\end{lstlisting}
\caption{Linear operator source code.} \label{fig:linear.m}
\end{figure}

@@

\subsubsection{$N$-ary operators}
All equations and boundary condition operators are defined
as operators \verb|{op1, op2, ... opN}|. 
Then, they are combined using the $N$-ary operator 
\verb|op = Join(op1, op2, ... opN)|, creating
the system operator \verb|op| (Figure \ref{fig:join.m}).

\begin{figure}[htpb]
\begin{lstlisting}
classdef Join < handle
properties
    ops;
end
methods
    function self = Join(operators)
        self.ops = operators;
        for k = 1:numel(self.ops)
        	op = self.ops{k}; 
        end
    end
    function r = res(self)
        r = cell(numel(self.ops), 1);
        for k = 1:numel(self.ops)
            r{k} = self.ops{k}.res();
        end
        r = cat(1, r{:});
    end
    function G = grad(self) 
        G = cell(numel(self.ops), 1);
        for k = 1:numel(self.ops)
            G{k} = self.ops{k}.grad();
        end
        G = cat(1, G{:});
    end
end
\end{lstlisting}
\caption{Join operator source code.} \label{fig:join.m}
\end{figure}

\subsection{Sample source code}
The following example shows how to define the following boundary-value problem:
\begin{align}
\Laplacian F(x,y) &= 1 & (x,y) &\in [0,1]^2 \\
F(0, y) = F(1, y) &= 0 & y &\in [0,1] \\
\deriv{F}{y}(x, 0) = \deriv{F}{y}(x, 1) &= 0 & x &\in [0,1]
\end{align}
and solve it using one Newton method step.
The source code is presented at Figure \ref{fig:sample.m}.
\begin{figure}[htpb]
\begin{lstlisting}
x = linspace(0, 1); 
y = linspace(0, 1);
g = Grid(x, y); % Grid definition
x_ = (x(1:end-1) + x(2:end)) * 0.5;
y_ = (y(1:end-1) + y(2:end)) * 0.5;
xi = x(2:end-1);
yi = y(2:end-1);
v = Variable(zeros(g.size)); 
F = Linear(g, v, speye(g.numel)); % Variable Definition
Fx = Deriv(Grid(x_, y), F, 1);
Fy = Deriv(Grid(x, y_), F, 2);
Fxx = Deriv(Grid(xi, yi), Fx, 1);
Fyy = Deriv(Grid(xi, yi), Fy, 2);
eq = {(Fxx + Fyy) - Const(1)};               % Laplacian(F) = 1
bnd = {Interp(Grid(0.0, y(2:end-1)), F); ... % F(0, y) = 0
	   Interp(Grid(1.0, y(2:end-1)), F); ... % F(1, y) = 0
       Interp(Grid(x, 0.0), Fy); ...         % dF/dy(x, 0) = 0
       Interp(Grid(x, 1.0), Fy)}		     % dF/dy(x, 1) = 0       
op = Join([eq; bnd]); % System operator
r = op.res()
G = op.grad()
v.update(-(G\r)) % Solve by Newton iteration
\end{lstlisting}
\caption{Sample problem source code.} \label{fig:sample.m}
\end{figure}

@@

For $\beta = 0$ (no electric field is applied), the steady-state solution is $\cU = 0$:
\begin{equation}\begin{array}{cccc}
\varPhi_0(r,\theta) = 0, &
C_0(r,\theta) = 1, &
\bV_0(r,\theta) = \bzero, &
P_0(r,\theta) = 0.
\end{array}\end{equation}
Assuming $\beta \ll 1$, the linearized equations and boundary conditions are
\begin{equation} \begin{array}{ccc}
\Laplacian \varPhi_1 = 0, &
\Laplacian C_1 = 0, &
\bLaplacian \bV_1 - \bnabla P_1 = \bzero.
\end{array}\end{equation}
The first-order terms $\bx_1$ are given by (see \cite{yariv2010migration} for full derivation)
\begin{equation} \begin{array}{cccc}
\varPhi_1 = \pars{\frac{1}{4r^2} - r}\cos\theta, &
C_1 = \frac{3}{4r^2} \cos\theta, &
\Psi_1 = \cU_1 \pars{\frac{1}{r} - r^2} \frac{\sin^2\theta}{2}, &
P_1 = 0,
\end{array} \end{equation}
and the linear velocity term is:
\begin{align} \label{eq:linear_velocity}
\bV_1 &= \cU_1 \pars{ \pars{-1 + r^{-3}}\cos\theta \cdot \brhat 
+ \pars{1 + r^{-3} / 2} \sin\theta \cdot \bthetahat } \\
\cU_1(\gamma) &= 2 \log \pars{\frac{1 + \gamma^{-\frac{1}{2}}}{2}}.
\end{align}
For $\gamma > 1$, the Debye layer has a positive space charge, 
corresponding to a negative particle charge -- 
yielding negative drift velocity, $\cU < 0$. 
For $\gamma < 1$, we have $\cU > 0$.

@@

\subsection{Ion-exchanger results}
For $\beta \ll 1$, the linear regime is dominant \eqref{eq:linear_velocity}.
When $\gamma \approx 1$, the cubic regime, where $\cU = O(\beta^3)$, 
dominates the linear one, even for weak electric fields. 
This phenomenon is confirmed both by the numerical
results and by asymptotic expansion of the nonlinear system (for $\alpha=0$).
The switch from the linear regime to the cubic regime occurs at 
an electric field value of $\beta = O(\sqrt{|\gamma - 1|})$.
The numerical results $\cU^h$, 
as well as the analytical ones \eqref{eq:cubic}, are given in Figure \ref{fig:IonExCubic}.
The numerical solver uses $N_r \times N_\theta = (400 \times 50)$, 
$R_{\max} = 10^7$ ($\Delta_r \approx 0.041$),
$K = 3$, $M = 5$ and $\beta = 10^{t}$ for $t = -2:0.1:0.5$.

@@

Since the leading order of the solution $\cU(\beta)$ is linear for $\beta \ll 1$, 
the numerical discretization error is $O(\beta)$ as well.
In order to reduce the discretization error of the numerical solution \eqref{eq:disc_zero_force}
and after verifying that the discretization error converges as $O(h^2)$,
we employ Richardson extrapolation to $\cU^h$ and $\cU^{2h}$:
\begin{equation}
\hat\cU = \frac{4}{3} \cU^h - \frac{1}{3} \cU^{2h}.
\end{equation}
The results of the extrapolation and $\hat\cU$ the analytical results 
are given at Figures \ref{fig:Electrophoresis1} and \ref{fig:Electrophoresis2}.
The numerical solver uses $R_{\max} = 10^2$,
$K = 3$, $M = 5$ and $\beta = 0.1 \cdot 2^{t}$ for $t = -2:0.5:5$.
The solver is applied to 2 grids of $N_r \times N_\theta = (128 \times 128)$ as the coarse grid 
(with $\Delta_r \approx 0.037$), and $N_r \times N_\theta = (256 \times 256)$ as the fine grid
(with $\Delta_r \approx 0.018$), in order to use Richardson extrapolation for $\hat\cU$.

@@

The computational grid is created to have fine resolution near the particle surface, while
approximating the far-away boundary conditions. The logarithmic discretization is motivated
by the solution of the linearized Stokes problem, where the largest gradients are near the particle
and the solution approaches the outer boundary conditions far away from the particle.
